What the hell is the title of this paper?

### 2018-05-14 NIPS 2018/main.tex
added MGVT app

Although this new model has the potential to capture much more complex distributions, it poses a major complication: the KL divergence between two mixture of Gaussians has no closed-form expression. To solve this issue, we rely on an upper bound to such quantity, so that negative ELBO still upper bounds the KL between the approximate and true posterior. Among the many upper bounds available in the literature, we adopt the one proposed in \cite{hershey2007approximating}, which we report here for the sake of completeness. We refer the reader to the original paper for the proof.
\begin{theorem}[\cite{hershey2007approximating}]
Let $p = \sum_{i} c_i^{(p)} f_i^{(p)}$ and $q = \sum_{j} c_j^{(q)} f_j^{(q)}$ be two mixture of Gaussian distributions, where $f_i^{(p)} = \mathcal{N}(\bm{\mu}_i^{(p)},\bm{\Sigma}_i^{(p)})$ denotes the i-th component of $p$, $c_i^{(p)}$ denotes its weight, and similarly for $q$. Introduce two vectors $\chi^{(1)}$ and $\chi^{(2)}$ such that $c_i^{(p)} = \sum_j \chi^{(2)}_{j,i}$ and $c_j^{(q)} = \sum_i \chi^{(1)}_{i,j}$. Then:
\begin{equation}
KL(p||q) \leq KL(\chi^{(2)}||\chi^{(1)}) + \sum_{i,j} \chi^{(2)}_{j,i}KL(f_i^{(p)}||f_j^{(q)})
\end{equation}
%\begin{equation}
%\chi^{(2)}_{i,j} = \frac{c_j^{(q)}\chi^{(1)}_{j,i}}{\sum_{i'} \chi^{(1)}%_{j,i'}},\ \ \ \ \chi^{(1)}_{j,i} = \frac{c_i^{(p)}\chi^{(2)}_{i,j}e^{-%KL(f_i^{(p)}||f_j^{(q)})}}{\sum_{j'}\chi^{(2)}_{i,j'}e^{-KL(f_i^{(p)}||f_{j'}%^{(q)})}}
%\end{equation}
\end{theorem}
Our new algorithm replaces the KL with the above-mentioned upper bound. Each time we require its value, we have to recompute the parameters $\chi^{(1)}$ and $\chi^{(2)}$ that tighten the bound. As shown in \cite{hershey2007approximating}, this can be achieved by a simple fixed-point procedure. Finally, both terms in the objective are now linear combinations of functions of the variational parameters of different components, and their gradients are easily derived from the ones of the Gaussian case. We report the derivation in App. \ref{}.

@@

Under Gaussian distributions, all quantities of interest for using Alg. \ref{alg:transfer} can be computed very easily. The KL divergence between the prior and approximate posterior can be computed in closed-form as:
\begin{equation}
KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \frac{1}{2}\left( \log\frac{\left\lvert \bm{\Sigma}_p \right\rvert}{\left\lvert \bm{\Sigma} \right\rvert} + \text{Tr}\left( \bm{\Sigma}_p^{-1}\bm{\Sigma} \right) + (\bm{\mu} - \bm{\mu}_p)^T \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p) -K \right)
\end{equation}
for $\bm{\xi} = (\bm{\mu},\bm{L})$ and $\bm{\Sigma} = \bm{LL}^T$. Its gradients with respect to the variational parameters are:
\begin{equation}
\nabla_{\bm{\mu}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p)
\end{equation}
\begin{equation}
\nabla_{\bm{L}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1}\bm{L} - (\bm{L}^{-1})^T
\end{equation}
Finally, the gradients w.r.t. the expected likelihood term of the variational objective \eqref{eq:elbo} can be computed using the reparameterization trick (e.g., \cite{hoffman2013stochastic,rezende2014stochastic}):
\begin{equation}
\nabla_{\bm{\mu}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation}
\begin{equation}
\nabla_{\bm{L}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \cdot \bm{v}^T \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation}

@@

>>>>

Although this new model has the potential to capture much more complex distributions, it poses a major complication: the KL divergence between two mixture of Gaussians has no closed-form expression. To solve this issue, we rely on an upper bound to such quantity, so that negative ELBO still upper bounds the KL between the approximate and true posterior. Among the many upper bounds available in the literature, we adopt the one proposed in \cite{hershey2007approximating}, which we report here for the sake of completeness. We refer the reader to the original paper for the proof.
\begin{theorem}[\cite{hershey2007approximating}]\label{th:ukl}
Let $p = \sum_{i} c_i^{(p)} f_i^{(p)}$ and $q = \sum_{j} c_j^{(q)} f_j^{(q)}$ be two mixture of Gaussian distributions, where $f_i^{(p)} = \mathcal{N}(\bm{\mu}_i^{(p)},\bm{\Sigma}_i^{(p)})$ denotes the i-th component of $p$, $c_i^{(p)}$ denotes its weight, and similarly for $q$. Introduce two vectors $\chi^{(1)}$ and $\chi^{(2)}$ such that $c_i^{(p)} = \sum_j \chi^{(2)}_{j,i}$ and $c_j^{(q)} = \sum_i \chi^{(1)}_{i,j}$. Then:
\begin{equation}
KL(p||q) \leq KL(\chi^{(2)}||\chi^{(1)}) + \sum_{i,j} \chi^{(2)}_{j,i}KL(f_i^{(p)}||f_j^{(q)})
\end{equation}
%\begin{equation}
%\chi^{(2)}_{i,j} = \frac{c_j^{(q)}\chi^{(1)}_{j,i}}{\sum_{i'} \chi^{(1)}%_{j,i'}},\ \ \ \ \chi^{(1)}_{j,i} = \frac{c_i^{(p)}\chi^{(2)}_{i,j}e^{-%KL(f_i^{(p)}||f_j^{(q)})}}{\sum_{j'}\chi^{(2)}_{i,j'}e^{-KL(f_i^{(p)}||f_{j'}%^{(q)})}}
%\end{equation}
\end{theorem}
Our new algorithm replaces the KL with the above-mentioned upper bound. Each time we require its value, we have to recompute the parameters $\chi^{(1)}$ and $\chi^{(2)}$ that tighten the bound. As shown in \cite{hershey2007approximating}, this can be achieved by a simple fixed-point procedure. Finally, both terms in the objective are now linear combinations of functions of the variational parameters of different components, and their gradients are easily derived from the ones of the Gaussian case. We report the derivation in App. \ref{}.

@@

Under Gaussian distributions, all quantities of interest for using Alg. \ref{alg:transfer} can be computed very easily. The KL divergence between the prior and approximate posterior can be computed in closed-form as:
\begin{equation}
KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \frac{1}{2}\left( \log\frac{\left\lvert \bm{\Sigma}_p \right\rvert}{\left\lvert \bm{\Sigma} \right\rvert} + \text{Tr}\left( \bm{\Sigma}_p^{-1}\bm{\Sigma} \right) + (\bm{\mu} - \bm{\mu}_p)^T \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p) -K \right)
\end{equation}
for $\bm{\xi} = (\bm{\mu},\bm{L})$ and $\bm{\Sigma} = \bm{LL}^T$. Its gradients with respect to the variational parameters are:
\begin{equation}\label{eq:klgradmu}
\nabla_{\bm{\mu}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1} (\bm{\mu} - \bm{\mu}_p)
\end{equation}
\begin{equation}\label{eq:klgradL}
\nabla_{\bm{L}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) = \bm{\Sigma}_p^{-1}\bm{L} - (\bm{L}^{-1})^T 
\end{equation}
Finally, the gradients w.r.t. the expected likelihood term of the variational objective \eqref{eq:elbo} can be computed using the reparameterization trick (e.g., \cite{hoffman2013stochastic,rezende2014stochastic}):
\begin{equation}\label{eq:tdgradmu}
\nabla_{\bm{\mu}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation} 
\begin{equation}\label{eq:tdgradL}
\nabla_{\bm{L}} \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{\mu},\bm{LL}^T)}\left[ ||B(\bm{w})||_{D}^2 \right] = \mathbb{E}_{\bm{v} \sim \mathcal{N}(\bm{0},\bm{I})}\left[ \nabla_{\bm{w}} ||B(\bm{w})||_{D}^2 \cdot \bm{v}^T \right]\ \ \text{for}\ \ \bm{w} = \bm{Lv} + \bm{\mu}
\end{equation}

@@

For the implementation of the Mixture of Gaussian Variational Transfer, we use the upper bound on the KL divergence between two Mixtures of Gaussians, as in Theorem\ref{th:ukl}, to obtain an upper bound on the negative ELBO in Equation\ref{eq:elbo}. Consider we have $C$ components for the posterior family $q_{\bm{\xi}}(\bm{w}) = \frac{1}{C}\sum_{i=1}^C\mathcal{N}(\bm{w}|\bm{\mu}_i,\bm{\Sigma}_i)$ and a prior distribution, constructed from the set of weights $\mathcal{W}_s = \left\lbrace \bm{w}_1,..., \bm{w}_{|\mathcal{W}_s|} \right\rbrace$ of the sources' optimal $Q$-functions, $p(\bm{w}) = \frac{1}{\abs{\mathcal{W}_s}}\sum_{j=1}^{|\mathcal{W}_s|}\mathcal{N}(\bm{w}|\bm{w}_j,\sigma_p^2\bm{I})$.

\begin{equation}\label{eq:ukl}
KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right) \leq KL(\chi^{(2)}||\chi^{(1)}) + \sum_{i=1}^C\sum_{j=1}^{|\mathcal{W}_s|} \chi_{j,i}^{(2)}KL(\mathcal{N}(\bm{w}|\bm{\mu}_i, \bm{\Sigma}_i)\ ||\ \mathcal{N}(\bm{w}|\bm{w}_j, \sigma_p^2\bm{I}))
\end{equation}

And substituting \ref{eq:ukl} in the negative ELBO in \ref{eq:elbo} we get the following upper bound.

\begin{equation}
\mathcal{L}(\bm{\xi}) \leq  \widetilde{\mathcal{L}}(\bm{\xi}) = \mathbb{E}_{\bm{w} \sim q_{\bm{\xi}}}\left[\norm{B(\bm{w})}_D^2\right] + KL(\chi^{(2)}||\chi^{(1)}) + \frac{\lambda}{N}\sum_{i=1}^C\sum_{j=1}^{|\mathcal{W}_s|} \chi_{j,i}^{(2)}KL(\mathcal{N}(\bm{w}|\bm{\mu}_i, \bm{\Sigma}_i)\ ||\ \mathcal{N}(\bm{w}|\bm{w}_j, \sigma_p^2\bm{I}))
\end{equation}

Finally, using this upper bound as objective of our optimization problem, we can then exploit the linearity of the expectation operator to obtain

\begin{equation}
\widetilde{\mathcal{L}}(\bm{\xi}) = \frac{1}{C}\sum_{i=1}^C \mathbb{E}_{\bm{w} \sim \mathcal{N}(\bm{w}|\bm{\mu}_i, \bm{\Sigma}_i)}\left[\norm{B(\bm{w})}_D^2\right] + KL(\chi^{(2)}||\chi^{(1)}) + \frac{\lambda}{N}\sum_{i=1}^C\sum_{j=1}^{|\mathcal{W}_s|} \chi_{j,i}^{(2)} KL(\mathcal{N}(\bm{w}|\bm{\mu}_i, \bm{\Sigma}_i)\ ||\ \mathcal{N}(\bm{w}|\bm{w}_j, \sigma_p^2\bm{I}))
\end{equation}

that is easily differentiable with respect to $\bm{\xi} = (\bm{\mu}_1,..., \bm{\mu}_C, \bm{\Sigma}_1,..., \bm{\Sigma}_C)$ using the Eq. \ref{eq:klgradmu}, \ref{eq:klgradL}, \ref{eq:tdgradmu}, \ref{eq:tdgradL}.
